### 2023-08-17 Data & Scripts/PCA.Rmd
Aug.17 pt.2
Desktop commit

---
title: "PCA"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: html_document
---

---

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
library("factoextra")
library("FactoMineR")
library("corrr")
library("ggcorrplot")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

## Preamble

This script will cover the pathway analysis. I've already performed a random permutation test in BORIS, but I would like to look at the individual level.

To do this, I've extracted the frequencies of one behavior following another (Head Down => Head Up) and will perform a pca to reduce the number of models I need to run.

That being said, I would, at most, need to run 9 models. However, I believe that certain transitions can be grouped together (e.g. transitions to 'alert' state).

To do this, I exported the behavioral sequences in BORIS, then calculated the frequencies of each transition, per individual, and compiled the 'PTWY' dataset.

```{r Open PTWY}
PCA<-read.csv("DATA.SR.csv", stringsAsFactors = T)
str(PCA)
```

Visualizing the data will be difficult in R, but the summaries and their visualizations can be found in Alex-Popescu-Thesis-2023\BORIS\Pathway Analysis\Flow Diagrams.

## Data distribution

```{r Data prep}
PCA.DATA<-PCA[1:81,c(2, 13, 31:34, 36:39, 41:44, 47, 50, 53, 56, 59, 62, 65, 68, 71, 74)]
PCA.DATA$HU_RATE<-PCA.DATA$HU_NUMBER_OF_BOUTS/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HD_RATE<-PCA.DATA$HD_NUMBER_OF_BOUTS/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$M_RATE<-PCA.DATA$M_NUMBER_OF_BOUTS/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HD.HDP_RATE<-PCA.DATA$NB_HD.HDP/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HD.HU_RATE<-PCA.DATA$NB_HD.HU/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HD.M_RATE<-PCA.DATA$NB_HD.M/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HDP.HU_RATE<-PCA.DATA$NB_HDP.HU/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HDP.M_RATE<-PCA.DATA$NB_HDP.M/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HU.HD_RATE<-PCA.DATA$NB_HU.HD/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HU.M_RATE<-PCA.DATA$NB_HU.M/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$M.HD_RATE<-PCA.DATA$NB_M.HD/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$M.HU_RATE<-PCA.DATA$NB_M.HU/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA<-PCA.DATA[1:81,c(1,4:6,8:10,12:15,25:36)]
```
```{r Distribution RAW}
PCA.DATA.LONG<-na.omit(pivot_longer(PCA.DATA, -ID))

ggplot(PCA.DATA.LONG
       , aes(x=value)) +
  geom_histogram() +
  geom_vline(data=PCA.DATA.LONG
       , aes(xintercept=mean(value), color="red")) +
  geom_vline(data=PCA.DATA.LONG
       , aes(xintercept=median(value), color="blue"))+
  facet_wrap(~name)
```

```{r Distribution LOG, include = F}
LOG<-as.data.frame(sapply(PCA.DATA[,-1], function(x) {log(x+0.01)}))
LOG.LONGER<-na.omit(pivot_longer(LOG, 1:22))

ggplot(LOG.LONGER
       , aes(x=value)) +
  geom_histogram() +
  geom_vline(data=LOG.LONGER
       , aes(xintercept=mean(value), color="red")) +
  geom_vline(data=LOG.LONGER
       , aes(xintercept=median(value), color="blue")) +
  facet_wrap(~name)
```

```{r Distribution ASIN SCALED}
SC.LOG<-as.data.frame(scale(LOG))
SC.LOG<-na.omit(SC.LOG)
SC.LOG.LONG<-pivot_longer(SC.LOG, 1:22)

ID = "ID."
ggplot(SC.LOG.LONG
       , aes(x=value)) +
  geom_histogram() +
  geom_vline(data=SC.LOG.LONG
       , aes(xintercept=mean(value), color="red")) +
  geom_vline(data=SC.LOG.LONG
       , aes(xintercept=median(value), color="blue")) +
  facet_wrap(~name)
```

## PCA

```{r PTWY pca}
#Computing correlation matrix
corr_matrix <- cor(na.omit(SC.LOG))

#Showing the correlation matrix
ggcorrplot(corr_matrix)

```

Pretty interesting. Redder = most positively correlated. Bluer = most negatively correlated. Let's apply the PCA.

```{r PCA scores}
#Computing the PCA
pca.data<-prcomp(SC.LOG)

#Show the results
summary(pca.data)
```

Pretty interesting. 
The first principal component explains 25.32% of the total variance. 
The second principal component explains 16.80% of the total variance. 
The third principal component explains 12.14% of the total variance.
The fourth principal component explains 9.74% of the total variance.
The fifth principal component explains 8.76% of the total variance.
The sixth principal component explains 7.54% of the total variance.
The seventh principal component explains 5.94% of the total variance.
The eighth principal component explains 4.86% of the total variance.

Let's visualize this with a scree plot.

```{r pca scree}
#The Scree plot to show how much of the variance each PC explains
fviz_eig(pca.data, addlabels = T, ylim = c(0,50))

```
The first eight PCs explain 91.11% of the variation!

## Graph of individuals

```{r PCA ind}
fviz_pca_ind(pca.data,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = cbPalette[1:3],
             repel = TRUE     # Avoid text overlapping
             )
```

## Graph of Variables

```{r PCA var}
fviz_pca_var(pca.data,
             col.var = "contrib", # Color by the quality of representation
             gradient.cols = cbPalette[1:3],
             repel = TRUE     # Avoid text overlapping
             )
```

## Eigenvalues

```{r PCA eig}
eig.val <- get_eigenvalue(pca.data)
eig.val

mean(eig.val$eigenvalue)
```

OK! Using the Kaiser-Guttman criterion, we end up retaining 8 PCs (mean eigenvalue = 0.4430, anything with an eigenvalue < 0.4430 is rejected). Let's look at how each factor fits into each principal component using their loadings.

```{r pca loadings}
-1*pca.data$rotation[,1:8]

#Criterion for "large" values
sqrt(1/ncol(SC.LOG)) #0.2132
```
Using the criterion for "large" results (=0.2132), we find that "peck rate", "HU Rate", "HU>HD Rate", and "HU.M Rate" do not fall into any PCs we've kept.

Should I remove them? Yes. Let's do that.

```{r remove columns}
colnames(SC.LOG)
SC.LOG2<-SC.LOG[,c(-10,-11,-19,-20)]
```

And now we re-run the PCA

## PCA2

```{r PTWY pca2}
#Computing correlation matrix
corr_matrix2 <- cor(na.omit(SC.LOG2))

#Showing the correlation matrix
ggcorrplot(corr_matrix2)

```

```{r PCA2 scores}
#Computing the PCA
pca.data2<-prcomp(SC.LOG2)

#Show the results
summary(pca.data2)
```

Nice!

The first 8 PCs explain a total of 92.11% of the variation, slightly more than the first PCA.

As before, let's plot a scree plot.

```{r PCA2 scree}
#The Scree plot to show how much of the variance each PC explains
fviz_eig(pca.data2, addlabels = T, ylim = c(0,50))

```

Generally, the PCs explain more variation that in the first PCA.

## Graph of individuals 2

```{r PCA ind}
fviz_pca_ind(pca.data2,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = cbPalette[1:3],
             repel = TRUE     # Avoid text overlapping
             )
```

## Graph of Variables 2

```{r PCA var}
fviz_pca_var(pca.data2,
             col.var = "contrib", # Color by the quality of representation
             gradient.cols = cbPalette[1:3],
             repel = TRUE     # Avoid text overlapping
             )
```

## Eigenvalues 2

```{r PCA eig}
eig.val <- get_eigenvalue(pca.data2)
eig.val

mean(eig.val$eigenvalue)
```

The mean eigenvalue is 0.5208. As a result, we keep 7 PCs. These PCs explain 87.17% of the total variation.

Let's look at the loadings again.

```{r pca loadings}
-1*pca.data2$rotation[,1:7]

#Criterion for "large" values
sqrt(1/ncol(SC.LOG2)) #0.2357
```

The magic number this time is 0.2357. All factors are included in the PCs. The next question is can I make this better? Is there any way for me to explain how these factors fit into each PC? Can I reduce the number of PCs?

RUN A MANOVA!!!