### 2023-08-17 Data & Scripts/PTWY.Rmd

---
title: "PTWY"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---
```{r Libraries}
library("lavaan")
library("lavaanPlot")
```

```{r Open DATA.SR}
PTWY<-read.csv("DATA.SR.csv", stringsAsFactors = T)
str(PTWY)
```
```{r PTWY Analysis}
PTWY.ENV <- '

# measurement model

ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TOTAL_FREQUENCY_OF_DISTURBANCES

# intercepts

CODED_ENV ~ 1
CODED_BAIT ~ 1
DECIMAL_TIME ~ 1
TOTAL_FREQUENCY_OF_DISTURBANCES ~ 1
'

PTWY.Fit.ENV <- sem(PTWY.ENV, data = PTWY)

summary(PTWY.Fit.ENV, fit.measures = TRUE, standardized = TRUE)
```

```{r PTWY.SOC Analysis}
PTWY.Mod.SOC <- '

# measurement model 
SOCIAL =~ CODED_GROUP_SIZE

# intercepts
CODED_GROUP_SIZE ~ 1
'

PTWY.Fit.SOC <- sem(PTWY.Mod.SOC, data = PTWY)

summary(PTWY.Fit.SOC, fit.measures = TRUE, standardized = TRUE)
```

```{r PTWY Analysis}
PTWY.Mod <- '

# measurement model

ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TEMPERATURE + TOTAL_FREQUENCY_OF_DISTURBANCES

VIGILANCE =~ PTWY_HDP.HU + PTWY_HD.HU + PTWY_M.HU

FORAGING =~ PTWY_HD.HDP + PTWY_HU.HD + PTWY_M.HD

# regressions

VIGILANCE ~ ENVIRONMENT + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE

FORAGING ~ ENVIRONMENT + VIGILANCE + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE

'

PTWY.Fit <- sem(PTWY.Mod, data = PTWY)

summary(PTWY.Fit, fit.measures = TRUE, standardized = TRUE)
```