### 2023-08-18 Data & Scripts/PTWY.Rmd

---
title: "PTWY"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---
```{r Libraries, include = F}
library("MANOVA.RM")
library("dplyr")
library("ggplot2")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

```{r Open DATA.SR}
PTWY.raw<-read.csv("DATA.SR.csv", stringsAsFactors = T)
str(PTWY.raw)
```

Nice, it opened as it should. Let's subset the dataset to only contain what we want.

```{r Subsetting PTWY}
PTWY<- PTWY.raw %>%
  dplyr::select(VIDEO_ID.,ID., DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_FREQUENCY_OF_DISTURBANCES, DISTURBANCE_FREQUENCY | starts_with("PTWY"))

```

Next, let's group together the transitions to vigilance, and transitions to vulnerability.

```{r PTWY group}
Vulnerable<-cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)
Vigilant<-cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)
# This does not work

```

Seems like it worked. Let's try fitting the model.

## Vulnerable

```{r MANOVA Vulnerable}
PTWY.vuln<-MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = PTWY, subject = "ID.", iter=1000)
summary(PTWY.vuln)
```

Interesting... While Sentinel presence and the interaction of sentinel presence and generalized environment had no significant effects, generalized environment had an effect on the frequency of transitions to vulnerability.

**Figure out if you need to recalculate PTWY data to only include those 4 transitions. **

Let's perform a pairwise comparison.

### Post Hoc

```{r PTWY CR Vuln}
simCI(PTWY.vuln, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "GENERALIZED_ENVIRONMENT")

```

HMMMM... Very weird... Now its not significant... Let's compute univariate comparisons.

```{r PTWY Vulnerable univar}
PTWY.HD.HDP<-MANOVA.wide(PTWY_HD.HDP~GENERALIZED_ENVIRONMENT, data = PTWY, iter = 1000)
PTWY.HU.HD<-MANOVA.wide(PTWY_HU.HD~GENERALIZED_ENVIRONMENT, data = PTWY, iter = 1000)

p.adjust(c(PTWY.HD.HDP$resampling[,1], PTWY.HU.HD$resampling[,1]), method = "bonferroni")
#using the parametric bootstrap Wald-type statistic

```

Very cool! Turns out that only the transitions from head up to head down is significantly affected by the generalized environment. This is only if we use the Wald-type statistic and not the modified ANOVA-type statistic.

### Plots

```{r PTWY plots}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
G.ENV.PTWY<-PTWY %>%
  summarySE(measurevar = "PTWY_HU.HD"
            , groupvars = "GENERALIZED_ENVIRONMENT") %>%
    ggplot(aes(x = GENERALIZED_ENVIRONMENT
                     , y = PTWY_HU.HD
                     , color = GENERALIZED_ENVIRONMENT))+
  geom_point(position = position_dodge(width=0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(PTWY_HU.HD-se)
                    , ymax=(PTWY_HU.HD+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Head Up to Head Down (%)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())

G.ENV.PTWY

```

## Vigilance

```{r MANOVA Vigilance}
PTWY.vigil<-MANOVA.wide(cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = PTWY, subject = "ID.", iter=1000)
summary(PTWY.vigil)
```

The results of this test suggest that no transitions to vigilance are significantly affected by either the presence of a sentinel or the generalized environment.

**Figure out what to do with factors such as bait presence, group size and disturbance frequency**

<!-- # Old code below -->
<!-- # ```{r PTWY Analysis} -->
<!-- # PTWY.ENV <- '  -->
<!-- #  -->
<!-- # # measurement model  -->
<!-- #  -->
<!-- # ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TOTAL_FREQUENCY_OF_DISTURBANCES  -->
<!-- #  -->
<!-- # # intercepts  -->
<!-- #  -->
<!-- # CODED_ENV ~ 1 -->
<!-- # CODED_BAIT ~ 1 -->
<!-- # DECIMAL_TIME ~ 1 -->
<!-- # TOTAL_FREQUENCY_OF_DISTURBANCES ~ 1 -->
<!-- # '  -->
<!-- #  -->
<!-- # PTWY.Fit.ENV <- sem(PTWY.ENV, data = PTWY)  -->
<!-- #  -->
<!-- # summary(PTWY.Fit.ENV, fit.measures = TRUE, standardized = TRUE) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r PTWY.SOC Analysis} -->
<!-- # PTWY.Mod.SOC <- '  -->
<!-- #  -->
<!-- # # measurement model  -->
<!-- # SOCIAL =~ CODED_GROUP_SIZE  -->
<!-- #  -->
<!-- # # intercepts -->
<!-- # CODED_GROUP_SIZE ~ 1 -->
<!-- # '  -->
<!-- #  -->
<!-- # PTWY.Fit.SOC <- sem(PTWY.Mod.SOC, data = PTWY)  -->
<!-- #  -->
<!-- # summary(PTWY.Fit.SOC, fit.measures = TRUE, standardized = TRUE) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r PTWY Analysis} -->
<!-- # PTWY.Mod <- '  -->
<!-- #  -->
<!-- # # measurement model  -->
<!-- #  -->
<!-- # ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TEMPERATURE + TOTAL_FREQUENCY_OF_DISTURBANCES  -->
<!-- #  -->
<!-- # VIGILANCE =~ PTWY_HDP.HU + PTWY_HD.HU + PTWY_M.HU  -->
<!-- #  -->
<!-- # FORAGING =~ PTWY_HD.HDP + PTWY_HU.HD + PTWY_M.HD  -->
<!-- #  -->
<!-- #   -->
<!-- #  -->
<!-- # # regressions  -->
<!-- #  -->
<!-- # VIGILANCE ~ ENVIRONMENT + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE -->
<!-- #  -->
<!-- # FORAGING ~ ENVIRONMENT + VIGILANCE + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE -->
<!-- #  -->
<!-- # '  -->
<!-- #  -->
<!-- # PTWY.Fit <- sem(PTWY.Mod, data = PTWY)  -->
<!-- #  -->
<!-- # summary(PTWY.Fit, fit.measures = TRUE, standardized = TRUE) -->
<!-- # ``` -->